---
title: "Ferula_RNAseq_Kallisto_DEGs_Drap_Oases_Plant6"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#After mapping reads in to kallisto, we want to find DEGs for Drap_Oases_Plant6.fasta, this fasta are from adding transcripts which are not available in drap assembly result in to drap assembly  
#consider facility as a batch effect which not true. Just keep the scrips to see the steps
#import data

```{r}
counts_All <- read.csv("~/Ferula_RNAseq_Analysis/Kallisto_out/combined_drap_oases_plant6_counts_kallisto.csv", header=T, row.names = "target_id")
head(counts_All) #for pooling two library of BR3 (they are not replication)
counts_All$BR3.2 <- counts_All$BR3 + counts_All$BR3.1
counts_All$BR3 <- NULL
counts_All$BR3.1 <- NULL
colnames(counts_All)[19] <- "BR3"
counts_All$BF3.2 <- counts_All$BF3 + counts_All$BF3.1
counts_All$BF3 <- NULL
counts_All$BF3.1 <- NULL
colnames(counts_All)[18] <- "BF3"
colnames(counts_All)[1] 
rownames(counts_All)
dim(counts_All) # 60134  18 60134  is num of transcripts in Reference
write.csv(counts_All, file="~/Ferula_RNAseq_Analysis/Kallisto_out/Ferula_RNAseq_combined_drap_oases_plant6_counts_kallisto.csv")
```

#filter based on read count, assign group, normalize, design matrix

```{r}
colSums(counts_All)
hist(colSums(counts_All,na.rm=TRUE))

#general threshold
colSums(counts_All,na.rm=TRUE) > 1000000 #just false for DF3
counts_All_nolow <- counts_All[,colSums(counts_All,na.rm=TRUE) > 1000000]
dim(counts_All_nolow) #612953 17 but I prefer to work with all 18 libraries
#sample description
samples <- data.frame(file=colnames(counts_All),
                      facility=factor(sub("(B|D|N)(S|F|L|R)(2|3|6)","\\1",colnames(counts_All))),
                      trt=factor(sub("(B|D|N)(S|F|L|R)(2|3|6)","\\2",colnames(counts_All))),
                    
                      genotype=factor(sub("(B|D|N)(S|F|L|R)(2|3|6)","\\3",colnames(counts_All)))) 
head(samples) 

#convert NA to zero
counts_All[is.na(counts_All)]<-0
# eliminating genes with low expression levels by retaining genes with > 10 reads in >= 3 samples
counts_All_small <-counts_All[rowSums(counts_All > 10) >= 3,] 
dim(counts_All_small) #36673  18 
dim(counts_All) #60134 18
write.csv(counts_All_small,file="~/Ferula_RNAseq_Analysis/Kallisto_out/Ferula_RNAseq_combined_drap_oases_plant6_counts_small_kalliso.csv") # use this for other analysis
```

```{r}
samples_small <- data.frame(file=colnames(counts_All_small),                     facility=factor(sub("(B|D|N)(S|F|L|R)(2|3|6)","\\1",colnames(counts_All_small))),
                      trt=factor(sub("(B|D|N)(S|F|L|R)(2|3|6)","\\2",colnames(counts_All_small))),
                    
                      genotype=factor(sub("(B|D|N)(S|F|L|R)(2|3|6)","\\3",colnames(counts_All_small)))) 
head(samples_small)
save(samples_small,file="~/Ferula_RNAseq_Analysis/Kallisto_out/Ferula_RNAse_combined_drap_oases_plant6_samples_small_kalliso.Rdata")
```

#assign group by combining all the experimental factors into one combined factor

```{r}
genotype<-levels(samples_small$genotype)
samples_small$group <- paste(samples_small$genotype,samples_small$trt,samples_small$facility,sep=".")
samples_small$genotype<-as.character(samples_small$genotype)
```
#edgeR

```{r}
library(edgeR)
dge <- DGEList(counts=counts_All_small, group=samples_small$group) 
length(colnames(counts_All_small)) # 18
dge<-calcNormFactors(dge, method = "TMM")
# look at the normalization factors
nrow(dge$samples) # 18 
hist(dge$samples[,3]) 
plot(log10(dge$sample[,"lib.size"]),dge$sample[,"norm.factors"]) 
#when we want to change the ref from first icon to others
samples_small$genotype <- as.factor(samples_small$genotype)
samples_small$genotype <- relevel(samples_small$genotype,ref="3") # plant3 reads including all three facilities
samples_small$trt <- as.factor(samples_small$trt)
samples_small$trt <- relevel(samples_small$trt,ref="R")
#design model for each factor (genotype & trt factors)
design1 <- model.matrix(~genotype+trt + facility, data=samples_small)
colnames(design1)
#First the overall dispersion
dge <- estimateGLMCommonDisp(dge,design1, verbose = T) #Disp =0.426 , BCV =0.653
dge <- estimateGLMTrendedDisp(dge,design1)
dge <- estimateGLMTagwiseDisp(dge,design1)
save(dge,file="~/Ferula_RNAseq_Analysis/Kallisto_drap_oases_plant6.dge.Rdata")
plotBCV(dge)
mds.dge <- plotMDS(dge, method = "bcv",labels = dge$samples$group)
```

#To find genes that are differentially expressed in gt 2 & 6 vs 3

```{r}
dge.fit <- glmFit(dge, design1)
#colnames(dge.fit)
dge.lrt <- glmLRT(dge.fit,coef = c("genotype2","genotype6"))
#the top 10 most differentially expressed genes
topTags(dge.lrt)
summary(decideTestsDGE(dge.lrt,p=0.05))
#Extract genes with a FDR < 0.01 (could also use 0.05)
DEgenes1 <- topTags(dge.lrt,n = Inf)$table[topTags(dge.lrt,n = Inf)$table$FDR<0.05,]
dim(DEgenes1) #7709    6
colnames(DEgenes1)
#head(DEgenes1)
#save 
write.csv(DEgenes1,file="~/Ferula_RNAseq_Analysis/Kallisto_out/Ferula_RNAseq_drap_oases_plant6_DEgenes1.csv")
#To find genes that are differentially expressed in gt 2 & 6 vs 3 seperately
dge.lrt.gt2 <- glmLRT(dge.fit,coef = c("genotype2"))
topTags(dge.lrt.gt2)
summary(decideTestsDGE(dge.lrt.gt2, p=0.05))
dge.lrt.gt6 <- glmLRT(dge.fit,coef = c("genotype6"))
topTags(dge.lrt.gt6)
summary(decideTestsDGE(dge.lrt.gt6, p=0.05))
DEGs1 <- data.frame ("gt2" = as.data.frame(summary(decideTestsDGE(dge.lrt.gt2, p=0.05)))$Freq,
                                     "gt6" = as.data.frame(summary(decideTestsDGE(dge.lrt.gt6, p=0.05)))$Freq)
rownames(DEGs1) <- c("down", "no", "up")
DEGs1 <- DEGs1[c("down", "up"),]
DEGs1
save(DEGs1, file="~/Ferula_RNAseq_Analysis/Kallisto_out/Ferula_RNAseq_drap_oases_plant6_DEGs1.Rdata")
```

```{r}
library(reshape2)
library(ggplot2)
DEGs1.melt <- melt(DEGs1)
DEGs1.melt$DE <- rep(c("down", "up"), 2)
colnames(DEGs1.melt) <- c("genotype", "number", "DE")
DEGs1.melt
# reorder: up 1st down 2nd 
DEGs1.melt$DE <- factor(DEGs1.melt$DE, levels = c("up", "down"))
DEGs1.melt <- DEGs1.melt[order(DEGs1.melt$DE),]
DEGs1.melt
DEGs1.melt$gt <- gsub("(X)(\\.)(S|L|F)(\\.)", "\\1",DEGs1.melt$genotype)

DEGs1.melt$trt <- gsub("(X)(\\.)(S|L|F)(\\.)", "\\2",DEGs1.melt$genotype)
```

#making ggplot for DEGs

```{r}
library(ggplot2)
p.DEGs1 <- ggplot(data = DEGs1.melt)
p.DEGs1 <- p.DEGs1 + geom_bar(mapping = aes(fill=DE, x = factor(DE), y = number) , stat= "identity")
p.DEGs1 <- p.DEGs1 + facet_grid(~genotype) 
p.DEGs1 <- p.DEGs1 + labs(y = "number of differentially expressed genes", x = "")
p.DEGs1
ggsave(p.DEGs1, file="~/Ferula_RNAseq_Analysis/Kallisto_out/Ferula_RNAseq_drap_oases_plant6_p.DEGs1.png")
```

#To find genes that are differentially expressed in trt S, F & L vs R 

```{r}
dge.lrt.trt <- glmLRT(dge.fit,coef = c("trtF","trtL", "trtS"))
#the top 10 most differentially expressed genes
topTags(dge.lrt.trt)
summary(decideTestsDGE(dge.lrt.trt,p=0.05))
#Extract genes with a FDR < 0.01 (could also use 0.05)
DEgenes2 <- topTags(dge.lrt.trt,n = Inf)$table[topTags(dge.lrt.trt,n = Inf)$table$FDR<0.05,]
write.csv(DEgenes2,file="~/Ferula_RNAseq_Analysis/Kallisto_out/Ferula_RNAseq_drap_oases_plant6_DEgenes2.csv")
dim(DEgenes2) #11778  7
colnames(DEgenes2)
head(DEgenes2)
```

##To find genes that are differentially expressed in trt S, F & L vs R seperately

```{r}
dge.lrt.trtF <- glmLRT(dge.fit,coef = c("trtF"))
topTags(dge.lrt.trtF)
summary(decideTestsDGE(dge.lrt.trtF, p=0.05))
dge.lrt.trtL <- glmLRT(dge.fit,coef = c("trtL"))
topTags(dge.lrt.trtL)
summary(decideTestsDGE(dge.lrt.trtL, p=0.05))
dge.lrt.trtS <- glmLRT(dge.fit,coef = c("trtS"))
topTags(dge.lrt.trtS)
summary(decideTestsDGE(dge.lrt.trtS, p=0.05))
```

```{r}
#For finding the expression of each genes for each tissues
a <- decideTestsDGE(dge.lrt.trtF, p=0.05)
#a[1:3]
#a[1:3,]
#a[1:3,1]
dim(a) #36673 1
b <- rownames(dge.lrt.trtF$coefficients)
c <- data.frame(b,a)
dge.lrt.trtL <- glmLRT(dge.fit,coef = c("trtL"))
d <- decideTestsDGE(dge.lrt.trtL, p=0.05)
e <- decideTestsDGE(dge.lrt.trtS, p=0.05)
c <- data.frame(b,a,d,e)
all_trts <- c
colnames(all_trts) <- c("Gene","Flower","Leaf","Stem")
write.table(all_trts, "~/Ferula_RNAseq_Analysis/Kallisto_out/all_trts_kallisto_drap_oases_plant6_DEgenes2.txt", col.names = T, row.names = T)
all_trtsv2 <- all_trts[-1]
```

#draw figure for DE genes across different tissues

```{r}
DEGs2 <- data.frame ("trtF" = as.data.frame(summary(decideTestsDGE(dge.lrt.trtF, p=0.05)))$Freq,
                                     "trtL" = as.data.frame(summary(decideTestsDGE(dge.lrt.trtL, p=0.05)))$Freq,
                                     "trtS"= as.data.frame( summary(decideTestsDGE(dge.lrt.trtS, p=0.05)))$Freq)
rownames(DEGs2) <- c("down", "no", "up")
DEGs2 <- DEGs2[c("down", "up"),]
DEGs2
save(DEGs2, file="~/Ferula_RNAseq_Analysis/Kallisto_out/Ferula_RNAseq_drap_oases_plant6_DEGs2.Rdata")
DEGs2.melt <- melt(DEGs2)
DEGs2.melt$DE <- rep(c("down", "up"), 3)
colnames(DEGs2.melt) <- c("genotype", "number", "DE")
DEGs2.melt
# reorder: up 1st down 2nd 
DEGs2.melt$DE <- factor(DEGs2.melt$DE, levels = c("up", "down"))
DEGs2.melt <- DEGs2.melt[order(DEGs2.melt$DE),]
DEGs2.melt
DEGs2.melt$gt <- gsub("(X)(\\.)(S|L|F)(\\.)", "\\1",DEGs2.melt$genotype)
DEGs2.melt$trt <- gsub("(X)(\\.)(S|L|F)(\\.)", "\\2",DEGs2.melt$genotype)
### making ggplot for DEGs
library(ggplot2)
p.DEGs2 <- ggplot(data = DEGs2.melt)
p.DEGs2 <- p.DEGs2 + geom_bar(mapping = aes(fill=DE, x = factor(DE), y = number) , stat= "identity")
p.DEGs2 <- p.DEGs2 + facet_grid(~genotype) 
p.DEGs2 <- p.DEGs2 + labs(y = "number of differentially expressed genes", x = "")
p.DEGs2
#save
ggsave(p.DEGs2, file="~/Ferula_RNAseq_Analysis/Kallisto_out/Ferula_RNAseq_drap_oases_plant6_p.DEGs2.png")
```

#create .txt file for creating gff3

```{r}
names <- rownames(DEgenes2[1:1000,])
names <- paste0("<",names)
names <- data.frame(names)
write.table(names, "top_transcripts_drap_oases_plant6.txt")

#for all DE genes
names2 <-rownames(DEgenes2)
names2 <- paste0("<",names2)
names2 <- data.frame(names2)
write.table(names2, "all_transcripts_drap_oases_plant6.txt")
```

